---
title: "Betke-Briana-ADA-homework-4"
author: "Briana Betke"
date: "5/8/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Homework 4
Starting with reading in the packages needed for the assignment:
```{r message=FALSE}
library(tidyverse)
library(patchwork) # for plotting
library(scales) # for log y axis scale
library(collapse) # for `fmode()` to find mode
library(BBmisc) # for %nin%
library(devtools)
library(tidycovid19)
```

run the function provided:
```{r message=FALSE}
merged <- download_merged_data(cached = TRUE)
merged <- merged %>%
  group_by(country, iso3c) %>%
  arrange(country, iso3c, date) %>%
  mutate(
    daily_confirmed = confirmed - lag(confirmed, n = 1),
    daily_deaths = deaths - lag(deaths, n = 1),
    daily_recovered = recovered - lag(recovered, n = 1)
  ) %>%
  mutate(
    daily_confirmed = replace_na(daily_confirmed, 0),
    daily_deaths = replace_na(daily_deaths, 0),
    daily_recovered = replace_na(daily_recovered, 0)
  ) %>%
  ungroup() %>%
  arrange(country, iso3c, date)

add_world1 <- merged %>%
  group_by(date) %>%
  arrange(date) %>%
  summarize(
    country = "World", iso3c = NA, confirmed = sum(confirmed, na.rm = TRUE),
    deaths = sum(deaths, na.rm = TRUE), recovered = sum(recovered, na.rm = TRUE),
    timestamp = fmode(timestamp)
  ) %>%
  mutate(
    daily_confirmed = confirmed - lag(confirmed, n = 1),
    daily_deaths = deaths - lag(deaths, n = 1),
    daily_recovered = recovered - lag(recovered, n = 1)
  ) %>%
  mutate(
    daily_confirmed = replace_na(daily_confirmed, 0),
    daily_deaths = replace_na(daily_deaths, 0),
    daily_recovered = replace_na(daily_recovered, 0)
  ) %>%
  ungroup() %>%
  arrange(country, iso3c, date)

add_world2 <- merged %>%
  group_by(country, iso3c) %>%
  summarize(
    population = fmode(population),
    land_area_skm = fmode(land_area_skm),
    timestamp = fmode(timestamp)
  ) %>%
  ungroup() %>%
  summarize(
    country = "World", iso3c = NA,
    population = sum(population, na.rm = TRUE),
    land_area_skm = sum(land_area_skm, na.rm = TRUE)
  ) %>%
  mutate(pop_density = population / land_area_skm)

add_world <- left_join(add_world1, add_world2, by = c("country", "iso3c"))
merged <- bind_rows(merged, add_world)

cv_data <- pivot_longer(merged,
  cols = c(
    "confirmed", "deaths", "recovered",
    "daily_confirmed", "daily_deaths", "daily_recovered"
  ),
  names_to = "variable", values_to = "cases"
) %>%
  arrange(country, variable, date) %>%
  rename(area = land_area_skm, density = pop_density) %>%
  mutate(rate = cases / population * 10^6) %>%
  ## new code to omit data before 2020-05-11
  filter(date < "2020-05-11")
## end of new code
```

Now the function:
```{r}
cv_summary <- function(d, country_list = "World",
                       plot = TRUE, facet = "country",
                       status = c("confirmed", "deaths", "recovered")) {

  # based on `wes_palettes()` color schemes GrandBudapest1, IsleofDogs1,
  # IsleofDogs2 from the {wesanderson} package
  my_palette <- c(
    "#5B1A18", "#FD6467", "#F1BB7B", "#D67236",
    "#0F0D0E", "#9986A5", "#79402E", "#CCBA72", "#D9D0D3", "#8D8680",
    "#EAD3BF", "#AA9486", "#B6854D", "#39312F", "#1C1718"
  )

  if (facet == "country") {
    fill <- "variable"
    n <- length(unique(d$variable)) / 2
    # need only half of unique # of variable (3)
  }

  if (facet == "variable") {
    fill <- "country"
    n <- length(country_list)
    # need number of countries
  }

  if ("All" %in% country_list) {
    country_list <- unique(d$country)
    country_list <- setdiff(country_list, "World")
  }

  if ("World" %in% country_list) {
    d <- d %>% filter(country %in% country_list)

    totals <- d %>%
      group_by(variable) %>%
      summarize(
        country = "World",
        cases = max(cases),
        population = max(population),
        area = max(area),
        density = max(density),
        rate = max(rate, na.rm = TRUE),
        on = max(date)
      ) %>%
      select(country, variable, cases, population, area, density, rate, on) %>%
      arrange(variable) %>%
      ungroup()
  }

  if ("World" %nin% country_list) {
    d <- d %>% filter(country %in% country_list)
    totals <- d %>%
      group_by(country, variable) %>%
      summarize(
        cases = max(cases),
        population = max(population),
        area = max(area),
        density = max(density),
        rate = max(rate, na.rm = TRUE),
        on = max(date),
        gdp_capita = fmode(gdp_capita),
        income = fmode(income),
        life_expectancy = fmode(life_expectancy),
        max_sd = max(soc_dist),
        max_mr = max(mov_rest)
      ) %>%
      select(
        country, variable, cases, population, area, density, rate,
        gdp_capita, income, life_expectancy, max_sd, max_mr, on
      ) %>%
      arrange(country, variable) %>%
      ungroup()
  }

  if (plot == TRUE) {
    cc <- filter(d, variable %in% status)
    cum_cases_plot <- ggplot(
      data = cc,
      # use the tidy evaluation pronoun .data to slice the chosen fill
      # variable from the data frame
      aes(
        x = date, y = cases + 1, color = .data[[fill]],
        fill = .data[[fill]]
      )
    ) +
      geom_point(size = 0.5) +
      geom_line() +
      # use the tidy evaluation pronoun .data to slice the chosen facet_wrap
      # variable from the data frame
      facet_wrap(~ .data[[facet]], ncol = 5) +
      xlab("Date") +
      ylab("Log Cumulative Cases") +
      scale_y_log10(
        breaks = trans_breaks("log10", function(x) 10^x),
        labels = trans_format("log10", math_format(10^.x))
      ) +
      scale_color_manual(
        aesthetics = c("color", "fill"),
        name = NULL, values = my_palette
      )

    dc <- filter(d, variable %in% paste0("daily_", status))
    daily_cases_plot <- ggplot(
      data = dc,
      aes(
        x = date, y = cases, color = .data[[fill]],
        fill = .data[[fill]]
      )
    ) +
      geom_point(size = 0.5) +
      geom_line() +
      facet_wrap(~ .data[[facet]], ncol = 5) +
      xlab("Date") +
      ylab("Daily Cases") +
      scale_color_manual(
        aesthetics = c("color", "fill"),
        name = NULL, values = my_palette
      )
  }

  if (plot == TRUE) {
    return(list(
      totals = totals,
      cum_cases_plot = cum_cases_plot,
      daily_cases_plot = daily_cases_plot
    ))
  } else {
    return(list(totals = totals))
  }
}
```

### Challenge 1 
Plot global data on confirmed coronavirus infections, deaths, and recoveries with the function. 
```{r}
results <- cv_summary(cv_data, country_list = "World")
results
```

### Challenge 2
Plot data on confirmed coronavirus infections, deaths, and recoveries for the G7 countries plus China, Russia, and Iran. Facet by country, then variable.
```{r}
# make a list of countries
countries <- list("United States", "United Kingdom", "Canada", "France", "Germany", "Italy", "Japan", "China", "Russia", "Iran")

# by country
cv_summary(cv_data, country_list = countries, facet = "country")

# by variable
cv_summary(cv_data, country_list = countries, facet = "variable")
```

### Challenge 3
Get the summary for all countries then filter for countries with a population over 1 million. Get the number of countires. 
```{r}
results2 <- cv_summary(cv_data, country_list = "All", plot = FALSE)
All <- results2$totals

# Tibble called d of countries with a population over 1 million
d <- filter(All, population > 1000000)

# I used the table function then got the lenght of the table. I know using the unique function in place of the table gives the same number.
length(table(d$country))
```
According to the output, there are 155 countries with populations greater than 1 million in this dataset. 

### Challenge 4
Create two datasets from d
```{r}
# Make the overall tibble
overall <- filter(d, variable == "confirmed")
head(overall, 3)

# Make the daily tibble
daily <- filter(d, variable == "daily_confirmed")
head(daily, 3)
```

Which 10 countries have exprienced the highest rate of confirmed cases?
```{r}
rate_All <- overall %>% arrange(desc(rate))
head(rate_All, 10)
```

Which 10 countries expericence the highest single day rate of confirmed cases?
```{r}
rate_Daily <- daily %>% arrange(desc(rate))
head(rate_Daily, 10)
```

### Challenge 5 
Running a linear model with overall rate as the response and density, population, gdp, and income as predictors.  

Visualizing the variables for overall:
```{r message=FALSE, warning=FALSE}
# Set up the histograpms with ggplot
rate_gg <- ggplot(overall, aes(x=rate)) + geom_histogram()
density_gg <- ggplot(overall, aes(x=density)) + geom_histogram()
pop_gg <- ggplot(overall, aes(x=population)) + geom_histogram()
gdp_gg <- ggplot(overall, aes(x=gdp_capita)) + geom_histogram()

# Make multiplot with patcwork since its already read in. 
(rate_gg | density_gg)/(pop_gg | gdp_gg)
```

Numeric variables with transformations:
```{r message=FALSE, warning=FALSE}
log_rate_gg <- ggplot(overall, aes(x=log(rate))) + geom_histogram()
log_density_gg <- ggplot(overall, aes(x=log(density))) + geom_histogram()
log_pop_gg <- ggplot(overall, aes(x=log(population))) + geom_histogram()
log_gdp_gg <- ggplot(overall, aes(x=log(gdp_capita))) + geom_histogram()

(log_rate_gg | log_density_gg)/(log_pop_gg | log_gdp_gg)
```

Run the model:
```{r}
overall_NA <- overall %>% drop_na()
# dataset without the value of zero in rate
overall_NA <- filter(overall_NA, rate > 0)
overall_NA$income <- as.factor(overall_NA$income)

mod <- lm(log(rate) ~ log(density) + log(population) + log(gdp_capita) + income, data = overall_NA)
summary(mod)
```
There is evidence, according to this model that density and gdp_capita have slopes that are significantly different than 0. 

### Challenge 6 
Stepwise selection with AIC find which model is preferred and do this also for the daily infection rates. 

```{r message=FALSE}
# Read in MASS and MuMIn
library(MASS)
library(MuMIn)

# running model selection from the model, called mod, defined in challenge 5
(s <- stepAIC(mod, scope = . ~ ., direction = "both"))
```

```{r}
# for the full model
r.squaredGLMM(mod)

# for the best model
best <- lm(log(rate) ~ log(density) + log(gdp_capita), data = overall_NA)
r.squaredGLMM(best)
```
Best model has lower R2 value than the full model.  

Now to repeat with daily data:
```{r}
# having trouble with running model with transformation
daily <- filter(daily, rate > 0)
daily$income <- as.factor(daily$income)

# Run the model
mod_daily <- lm(log(rate) ~ log(density) + log(population) + log(gdp_capita) + income, data = daily)
summary(mod_daily)
```

Variable selection
```{r}
(s <- stepAIC(mod_daily, scope = . ~ ., direction = "both"))
```
Pseudo r squared
```{r}
# for the full model
r.squaredGLMM(mod_daily)

# for the best model
best <- lm(log(rate) ~ log(density) + log(gdp_capita), data = daily)
r.squaredGLMM(best)
```
The output for daily appears to be the same as the overall. I am not sure this is correct.

### Challenge 7 
Add social distancing and movement restriction to the selected model from challenge 6.  

First model with both variables added.
```{r}
daily$max_sd_log1 <- log(daily$max_sd + 1)
daily$max_mr_log1 <- log(daily$max_mr + 1)

best_plus_all <- lm(log(rate) ~ log(density) + log(gdp_capita) + max_sd_log1 + max_mr_log1, data = daily)
summary(best_plus_all)
```

Second model without movement restriction:
```{r}
best_plus_sd <- lm(log(rate) ~ log(density) + log(gdp_capita) + max_sd_log1, data = daily)
summary(best_plus_sd)
```

Third model without social distancing:
```{r}
best_plus_mr <- lm(log(rate) ~ log(density) + log(gdp_capita) + max_mr_log1, data = daily)
summary(best_plus_mr)
```

Look at AICc table:
```{r message=FALSE}
library(AICcmodavg) 
aic_table <- aictab(list(best_plus_all, best_plus_sd, best_plus_mr),
  modnames = c("best_plus_all", "best_plus_sd", "best_plus_mr"))
aic_table
```
When looking at AICc, it appears that having one or the other, does improve the model. 

### Challenge 8 

Set up the dataset:
```{r}
dat <- cv_data %>% filter(population > 1000000 & rate > 0 & variable == "daily_confirmed")
```

On to the models!
```{r message=FALSE}
library(lme4) # for the lmer function
# Full model
m1 <- lmer(data = dat, log(rate) ~ log(density) + log(gdp_capita) + soc_dist + mov_rest + (1|country) + (1|date),   REML = FALSE)

# wihtout mov_rest
m2 <- lmer(data = dat, log(rate) ~ log(density) + log(gdp_capita) + soc_dist + (1|country) + (1|date),   REML = FALSE)

# without soc_dist
m3 <- lmer(data = dat, log(rate) ~ log(density) + log(gdp_capita) + (1|country) + (1|date),   REML = FALSE)

# without density
m4 <- lmer(data = dat, log(rate) ~ log(gdp_capita) + soc_dist + mov_rest + (1|country) + (1|date),   REML = FALSE)

# without gdp
m5 <- lmer(data = dat, log(rate) ~ log(density) + soc_dist + mov_rest + (1|country) + (1|date),   REML = FALSE)

# Null model with just random effects
null <- lmer(data = dat, log(rate) ~ (1|country) + (1|date), REML = FALSE)
```

Create and AIC table of the models. Which one is the best model? What is the pseudo r value of that model?
```{r message=FALSE}
(aic_table <- aictab(list(m1, m2, m3, m4, m5, null),
  modnames = c("m1", "m2", "m3", "m4", "m5", "null")))

r.squaredGLMM(m1)
```
According to the table, the full model has the lowest AICc and the highest Akaike weight of 1. 



